Creating the ``r params$package_name`` R package

### 2025-08-23 create-FSHybridPLS.Rmd
eigen and linear equivalent

```{r}

# --- Example Usage ---

# Example 1: All matrices are identical
mat1 <- matrix(c(1, 2, 3, 4), nrow = 2)
mat2 <- matrix(c(1, 2, 3, 4), nrow = 2)
mat3 <- matrix(c(1, 2, 3, 4), nrow = 2)
identical_list <- list(mat1, mat2, mat3)
print("Test Case 1: All matrices identical")
print(are_all_gram_matrices_identical(identical_list)) # Expected: TRUE

# Example 2: Matrices with slight floating-point differences (within tolerance)
mat4 <- matrix(c(1.000000000000001, 2, 3, 4), nrow = 2)
mat5 <- matrix(c(1, 2, 3, 4), nrow = 2)
tolerance_list <- list(mat4, mat5)
print("Test Case 2: Matrices with slight floating-point differences")
print(are_all_gram_matrices_identical(tolerance_list)) # Expected: TRUE

# Example 3: Matrices are not identical
mat6 <- matrix(c(1, 2, 3, 5), nrow = 2) # Different value
non_identical_list <- list(mat1, mat6, mat3)
print("Test Case 3: Matrices are not identical")
print(are_all_gram_matrices_identical(non_identical_list)) # Expected: FALSE

# Example 4: List with different dimensions
mat7 <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
diff_dim_list <- list(mat1, mat7)
print("Test Case 4: Matrices with different dimensions")
print(are_all_gram_matrices_identical(diff_dim_list)) # Expected: FALSE (all.equal will catch this)

# Example 5: Empty list
empty_list <- list()
print("Test Case 5: Empty list")
print(are_all_gram_matrices_identical(empty_list)) # Expected: TRUE

# Example 6: List with a single matrix
single_matrix_list <- list(mat1)
print("Test Case 6: Single matrix list")
print(are_all_gram_matrices_identical(single_matrix_list)) # Expected: TRUE

# Example 7: List containing non-matrix elements (will stop with an error)
# non_matrix_list <- list(mat1, "not a matrix", mat3)
# print("Test Case 7: List with non-matrix element (will throw error)")
# print(are_all_gram_matrices_identical(non_matrix_list))
```

@@

```{r}
#' Construct a Single-Sample Predictor Hybrid Object from Coefficients
#'
#' Reconstructs a \code{predictor_hybrid} object representing one observation, using a numeric
#' coefficient vector. This alternative constructor maps the coefficients back into their functional
#' and scalar predictor representations based on the structure of a template \code{predictor_hybrid} object.
#'
#' @param format A \code{predictor_hybrid} object that provides the structure and basis information.
#' @param coef A numeric vector containing coefficients for both functional and scalar predictors.
#'
#' @return A \code{predictor_hybrid} object with updated \code{functional_list}, \code{Z}, and \code{n_sample = 1}.
#'
#'
#' @export
predictor_hybrid_from_coef <- function(format, coef){
  M <- dim(format$gram_list[[1]])[2] # number of basis functions
  K <- length(format$functional_list) # number of functional predictors
  for (ii in 1:K){
    format$functional_list[[ii]] <- fd(
      coef = as.matrix(coef[((ii-1)*M + 1):(ii*M)]),
      basisobj = format$functional_list[[ii]]$basis
    )
  }
  format$Z <- t(as.matrix(coef[(K * M + 1):length(coef)]))
  format$n_sample <- 1
  return(format)
}
```

@@

@@

```{r}
#' Extract a single observation from a predictor_hybrid object
#'
#' @param W A predictor_hybrid object with multiple samples.
#' @param i Integer index of the sample to extract.
#' @return A single-sample predictor_hybrid object.
subset_predictor_hybrid <- function(W, i) {
  new_Z <- matrix(W$Z[i, ], nrow = 1)
  new_functional_list <- lapply(W$functional_list, function(fdobj) {
    fd(coef = matrix(coef(fdobj)[, i], ncol = 1), basisobj = fdobj$basis)
  })
  predictor_hybrid(Z = new_Z, functional_list = new_functional_list)
}

@@

  xi_star <- t(invL) %*% e      # xi_star solves t(L) xi = e
  xi_hat <- predictor_hybrid_from_coef(format = W, coef = xi_star)
  return(list(
    xi_hat = xi_hat,
    E = E,
    V_star = V_star,
    eigen_val = eigen_result$values[1]
  ))
}
```

@@

    # linear system
    R <- W$gram_list[[j]] + lambda[j] * W$gram_deriv2_list[[j]]
    U_cholesky <- chol(R)
    L_cholesky <- t(U_cholesky)
    intermediate <- forwardsolve(L_cholesky, u[[j]])
    gamma[[j]] <- backsolve(U_cholesky, intermediate)

     #q
    q <- q + sum( u[[j]] * gamma[[j]] )
  }
  d_vec <- c(do.call(c, gamma), v)  # or: unlist(d)
  ## step 2: squared norm
  d_vec <- d_vec/ sqrt(q)
  return(d_vec)
}
```

@@

n_sample <- 100
n_scalar <- 2
n_basis <- 5
n_functional <- 2

@@

```{r}
residualize_y <- function(y, rho){
  nu <- get_nu(y, rho)
  y_next <- y - nu * rho
  return(y_next)
}
```

output of the residualize_y and rho must be orthogonal.
```{r}
#sum(residualize_y(y, rho )*rho) #should output a very small number
```

@@

```{r}
residualize_predictor <- function(W, rho, delta){
  n <- length(rho)
  W_res <- W
  for (i in 1:n){
    W_res_i <-subtr.predictor_hybrid(subset_predictor_hybrid(W, i), delta,  rho[i]) 
    W_res <- replace_obs_hybrid(W_res, i, W_res_i)
  }
  return(W_res)
}
```

```{r}
delta <- get_delta(W,rho)
res <- residualize_predictor(W, rho, delta)
zero<-get_delta(res, rho)
```

@@

  # matrices in the statement of Proposition 4
  J_star <-as.matrix(get_jacobian_hybrid(W))
  d_hat_star <- t( #basis coefficient of hybrid regression coefficient
    solve(
      t((norm(rho, type="2"))^2 * J_star),
      t(rho_t_C_star_J_star)
      )
    )
  delta_hat <- hybrid_from_coef(W, d_hat_star) #hybrid regression coefficient
  return(delta_hat)
  }
```

```{r}
###########################################
)

setMethod("fitted_value",
          signature(
            "predictor_hybrid",
            "vector"
          ),
####################################################
function(regcoef, pls_score) {
  for (ii in 1 : (regcoef@n_functional) ){
    regcoef@functional_list[[ii]] <- fd(
        coef = coef(regcoef@functional_list[[ii]]) %*% matrix(pls_score, nrow=1),
        basisobj = regcoef@functional_list[[ii]]$basis
      )}
  regcoef@Z <-  pls_score %*% regcoef@Z
  return(regcoef)
  }
####################################################
)

@@

>>>>

@@

```{r}
#' Construct a Single-Sample Predictor Hybrid Object from Coefficients
#'
#' Reconstructs a \code{predictor_hybrid} object representing one observation, using a numeric
#' coefficient vector. This alternative constructor maps the coefficients back into their functional
#' and scalar predictor representations based on the structure of a template \code{predictor_hybrid} object.
#'
#' @param format A \code{predictor_hybrid} object that provides the structure and basis information.
#' @param coef A numeric vector containing coefficients for both functional and scalar predictors.
#'
#' @return A \code{predictor_hybrid} object with updated \code{functional_list}, \code{Z}, and \code{n_sample = 1}.
#'
#'
#' @export
predictor_hybrid_from_coef <- function(format, coef){
  M <- format$n_basis_list[1] # number of basis functions
  K <- format$n_functional # number of functional predictors
  for (ii in 1:K){
    format$functional_list[[ii]] <- fd(
      coef = as.matrix(coef[((ii-1)*M + 1):(ii*M)]),
      basisobj = format$functional_list[[ii]]$basis
    )
  }
  format$Z <- t(as.matrix(coef[(K * M + 1):length(coef)]))
  format$n_sample <- 1
  return(format)
}
```

@@

```{r}
#' Inner product between two predictor_hybrid objects (with broadcasting)
#'
#' Computes the inner product between two `predictor_hybrid` objects,
#' including both functional and scalar components. Supports broadcasting
#' when one of the inputs has a single observation.
#'
#' @param xi_1 A `predictor_hybrid` object.
#' @param xi_2 Another `predictor_hybrid` object. If missing, defaults to `xi_1`.
#'
#' @return A numeric vector of inner products, or a scalar if both inputs contain a single observation.
#' @export
inprod_pen.predictor_hybrid <- function(xi_1, xi_2 = NULL, lambda) {

  inprod <- inprod.predictor_hybrid(xi_1, xi_2)
  # Prepare components

  f1 <- xi_1$functional_list

  f2 <- xi_2$functional_list
  for (j in 1:xi_1$n_functional){
    inprod <- inprod + lambda[j] * fda::inprod(
        fdobj1 = f1[[j]], 
        fdobj2 = f2[[j]], 
        Lfdobj1 = 2,
        Lfdobj2 = 2
        )

    }

  # Combine results
   return(as.numeric(inprod))
}
```

@@

```{r}
#' Extract a single observation from a predictor_hybrid object
#'
#' @param W A predictor_hybrid object with multiple samples.
#' @param i Integer index of the sample to extract.
#' @return A single-sample predictor_hybrid object.
subset_predictor_hybrid <- function(W, i) {
  new_Z <- matrix(W$Z[i, ], nrow = 1)
  new_functional_list <- lapply(W$functional_list, function(fdobj) {
    fd(coef = matrix(coef(fdobj)[, i], ncol = 1), basisobj = fdobj$basis)
  })
  new_predictor <- predictor_hybrid(Z = new_Z, functional_list = new_functional_list)
  return(new_predictor)
}
```

@@

  xi_star <- t(invL) %*% e      # xi_star solves t(L) xi = e
  xi_hat <- predictor_hybrid_from_coef(format = W, coef = xi_star)
  return(xi_hat)
}
```

@@

    # linear system
    R <- W$gram_list[[j]] + lambda[j] * W$gram_deriv2_list[[j]]
    gamma[[j]] <- solve(R, u[[j]])

     #q
    q <- q + sum( gamma[[j]] * ( R %*% gamma[[j]] ))
  }
  d_vec <- c(do.call(c, gamma), v)  # or: unlist(d)
  ## step 2: squared norm
  d_vec <- d_vec/ sqrt(q)
  xi_hat <- predictor_hybrid_from_coef(format = W, coef = d_vec)
  return(xi_hat)
}
```

@@

n_sample <- 100
n_scalar <- 2
n_basis <- 20
n_functional <- 2

@@

```{r}
residualize_y <- function(y, rho, nu){
  y_next <- y - nu * rho
  return(y_next)
}
```

@@

```{r}
residualize_predictor <- function(W, rho, delta){
  n <- length(rho)
  W_res <- W
  for (i in 1:n){
    W_i <- subset_predictor_hybrid(W, i)
    W_res_i <-subtr.predictor_hybrid(W_i, delta,  rho[i]) 
    W_res <- replace_obs_hybrid(W_res, i, W_res_i)
  }
  return(W_res)
}
```

@@

  # matrices in the statement of Proposition 4
  J_star <-as.matrix(get_jacobian_hybrid(W))
  d_hat_star <- t( #basis coefficient of hybrid regression coefficientdelta
    solve(
      t((norm(rho, type="2"))^2 * J_star),
      t(rho_t_C_star_J_star)
      )
    )
  delta_hat <- hybrid_from_coef(W, d_hat_star) #hybrid regression coefficient
  return(delta_hat)
  }
```

# main algorithm

```{r eval=FALSE, include=FALSE}
#' @export
fit.hybridPLS <- function(W, y, n_iter, lambda) {
  # 1. Initialize storage
  W_now <- rho <- xi <- delta <- nu  <- sigma <- eta <- list()

  # 3. Initialize residuals
  W_now[[1]] <- W
  y_now <- y
  final_succesful_iteration <- 0
  constr_mat <- get_constraint_matrix_hybrid(W, lambda)
  constr_mat_chol <- t(chol(constr_mat))
  for (l in 1:n_iter) {
    cat(paste(l, "th component", "\n"))

    xi[[l]]  <- get_xi_hat_linear_pen(W_now[[l]], y_now, lambda) # PLS direction
     rho[[l]] <- inprod.predictor_hybrid(W_now[[l]], xi[[l]]) #PLS score
    delta[[l]] <- get_delta(W_now[[l]], rho[[l]]);  
    W_now[[l + 1]] <- residualize_predictor(W_now[[l]], rho[[l]], delta[[l]]) # predictor residual
    nu[[l]] <- get_nu(y_now, rho[[l]]); y_now <- residualize_y(y_now, rho[[l]], nu[[l]]) # response residual

    # Step 3: Orthogonalize and update prediction
    #sigma[[l]] <- xi[[l]]
    #if (l == 1) {
    #  eta[[l]] <- scalar_mul(sigma[[l]], nu[[l]])
    #} else {
    #  for (u in 1:(l - 1)) {
    #    sigma[[l]] <- subtr(sigma[[l]], sigma[[u]], hybrid_inner_prod(delta[[u]], xi[[l]]))
    #  }
    #  eta[[l]] <- add(eta[[l - 1]], sigma[[l]], nu[[l]])
    #}
  }
  return (list(
    rho = rho,
    xi = xi,
    W = W_now
  ))
}

@@

```{r eval=FALSE, include=FALSE}
#' @export
fit.hybridPLS_eigen <- function(W, y, n_iter, lambda) {
  # 1. Initialize storage
  W_now <- rho <- xi <- delta <- nu  <- sigma <- eta <- list()

  # 3. Initialize residuals
  W_now[[1]] <- W
  y_now <- y
  final_succesful_iteration <- 0
  constr_mat <- get_constraint_matrix_hybrid(W, lambda)
  constr_mat_chol <- t(chol(constr_mat))
  for (l in 1:n_iter) {
    cat(paste(l, "th component", "\n"))

    #xi[[l]]  <- get_xi_hat_linear_pen(W_now[[l]], y_now, lambda) # PLS direction
    xi[[l]]  <- get_pls_comp(W_now[[l]], y_now, constr_mat_chol) # PLS direction
    rho[[l]] <- inprod.predictor_hybrid(W_now[[l]], xi[[l]]) #PLS score
    delta[[l]] <- get_delta(W_now[[l]], rho[[l]]);  
    W_now[[l + 1]] <- residualize_predictor(W_now[[l]], rho[[l]], delta[[l]]) # predictor residual
    nu[[l]] <- get_nu(y_now, rho[[l]]); y_now <- residualize_y(y_now, rho[[l]], nu[[l]]) # response residual

    # Step 3: Orthogonalize and update prediction
    #sigma[[l]] <- xi[[l]]
    #if (l == 1) {
    #  eta[[l]] <- scalar_mul(sigma[[l]], nu[[l]])
    #} else {
    #  for (u in 1:(l - 1)) {
    #    sigma[[l]] <- subtr(sigma[[l]], sigma[[u]], hybrid_inner_prod(delta[[u]], xi[[l]]))
    #  }
    #  eta[[l]] <- add(eta[[l - 1]], sigma[[l]], nu[[l]])
    #}
  }
  return (list(
    rho = rho,
    xi = xi,
    W = W_now
  ))
}

```{r}
inprod.predictor_hybrid(res$W[[5]], res$xi[[3]])

sum(y * inprod.predictor_hybrid(res$W[[5]], res$xi[[3]]))

inprod_pen.predictor_hybrid(res$xi[[1]], res$xi[[8]], lambda) - inprod_pen.predictor_hybrid(res_eigen$xi[[1]], res_eigen$xi[[8]], lambda)
inprod_pen.predictor_hybrid(res$xi[[1]], res$xi[[8]], 10*lambda) - inprod_pen.predictor_hybrid(res_eigen$xi[[1]], res_eigen$xi[[8]], 10*lambda)
inprod_pen.predictor_hybrid(res$xi[[1]], res$xi[[8]], lambda/10) - inprod_pen.predictor_hybrid(res_eigen$xi[[1]], res_eigen$xi[[8]], lambda/10)
sum(y * inprod.predictor_hybrid(res$W[[8]], res$xi[[1]])) - sum(y * inprod.predictor_hybrid(res_eigen$W[[8]], res_eigen$xi[[1]]))

inprod_pen.predictor_hybrid(res$xi[[3]], res$xi[[4]], lambda) - inprod_pen.predictor_hybrid(res_eigen$xi[[3]], res_eigen$xi[[4]], lambda)
inprod_pen.predictor_hybrid(res$xi[[3]], res$xi[[4]], 10*lambda) - inprod_pen.predictor_hybrid(res_eigen$xi[[3]], res_eigen$xi[[4]], 10*lambda)
inprod_pen.predictor_hybrid(res$xi[[3]], res$xi[[4]], lambda/10)
sum(y * inprod.predictor_hybrid(res$W[[3]], res$xi[[1]]))

inprod_pen.predictor_hybrid(res$xi[[5]], res$xi[[10]], lambda)
inprod_pen.predictor_hybrid(res$xi[[5]], res$xi[[10]], 10*lambda)
inprod_pen.predictor_hybrid(res$xi[[5]], res$xi[[10]], lambda/10)
sum(y * inprod.predictor_hybrid(res$W[[5]], res$xi[[1]]))

inprod_pen.predictor_hybrid(res$xi[[4]], res$xi[[4]], lambda)
inprod_pen.predictor_hybrid(res$xi[[4]], res$xi[[4]], lambda/10)
inprod_pen.predictor_hybrid(res$xi[[4]], res$xi[[4]], 10*lambda )

```{r}
lambda <- c(0.5,0.5)
L=8
res <- fit.hybridPLS(W,y,L, lambda)
res_eigen <- fit.hybridPLS_eigen(W,y,L, lambda)
```
